---
title: "Week 5 Assignment - Tidying and Transforming Data"
author: "John Mazon"
date: "9/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### MYSQL Database integration with R

The MYSQL database table created for this assignment contains arrival delays for two airlines across five destinations

#### Additional packages installed
Please note below additional packages installed. The formattable package is used to transform vectors and data frames into more readable and impactful tabular formats. 
Package 'conflicted' solved an issue with the rename func as it pertains to knit/html.

install.packages("formattable")
install.packages("conflicted")

```{r}
library(tidyverse)
library(RMySQL)
library(formattable)

```


```{r}
##DBconnect was utilized in this assignment to authenticate/pull in the flight delay information into R
my.database = dbConnect(MySQL(), user='root', password = 'Password1', dbname='flightsandstatus', host='localhost')
dbListTables(my.database)

```
Using fetch function we are able to select all the row from our flightsandstatus database [table = airlines_info1]
```{r}
airlines_status <- fetch(dbSendQuery(my.database, "SELECT * FROM airlines_info1 ORDER BY ID"))
dim(airlines_status)
colnames(airlines_status)
summary(airlines_status)
```

Experimented with the tibble func below for the first time. Tibbles have a refined print method that shows only the first 10 rows, and all the columns that fit on screen. This makes it much easier to work with large data.
```{r}
airline_statusinfo <- as_tibble(airlines_status)
airline_statusinfo

```
Note the use of the conflicted library below was helpful in tackling a difficult issue of knitting to html with the rename func used in our R code. Similarly from assignment 2, it is ideal for us to rename often time certain data which is not easy to understand/or to better the appearance of certain data such as columns. Also note the dplyr::rename being utilized in this scenario.
```{r}
library(conflicted)
library(dplyr)

airlinestatusinfo9 <- dplyr::rename(airline_statusinfo,"Los Angeles"="cities_losangeles",
          "Phoenix" = "cities_phoenix",
          "San Diego" = "cities_sandiego",
          "San Francisco" = "cities_sanfrancisco",
          "Seattle" = "cities_seattle"
)

print(airlinestatusinfo9)

```
## Data Tidying 
Note below we are using the gather, select and spread func to tidy/better organize the flights info to later on be able to manipulate and do calculations. We were encouraged to use a 'wide' structure so that we are later able to transform it.


```{r}
airline_statusinfo8 <- airlinestatusinfo9 %>% 
  gather(City, count, -ID, -Airline, -Status) %>% 
  select(Airline, Status, City, count) %>% 
  spread(Status, count)

  
print(airline_statusinfo8)


```

The formattable package is used to transform vectors and data frames into more readable and impactful tabular formats. 
```{r}
formattable(airline_statusinfo8)

```
## Data Analysis 
### Perform analysis to compare the arrival delays for the two airlines.
```{r}

airline_statusinfo7 <- airline_statusinfo8 %>%
  dplyr::mutate(
         ontime_rate = `on time` / (delayed + `on time`),
         delay_rate = delayed / (delayed + `on time`),
         flights = delayed + `on time`,
         percent_on_time = round(`on time`/flights*100, 2),
         
         
  )
formattable(airline_statusinfo7)

```
Using ggplot2 to gather the delay information based on airline
```{r}
library(ggplot2)
Alaska_delays<-dplyr::filter(airline_statusinfo7, Airline=="ALASKA")
Alaska_delays

AMWEST_delays<-dplyr::filter(airline_statusinfo7, Airline=="AM WEST")
AMWEST_delays

```

```{r}
compare_data <- airline_statusinfo7 %>% 
  group_by(Airline) %>%
  dplyr::summarize(OnTime = sum(`on time`), Delayed = sum(delayed), 
            PercentOnTime = round(OnTime/sum(OnTime,Delayed),2))%>%
  select(Airline, PercentOnTime)
compare_data %>% dplyr::arrange(desc(PercentOnTime))



```

```{r}
 ggplot(airline_statusinfo7, aes(x = Airline, y=delay_rate, fill = City)) +  
          geom_bar(stat="identity",position="dodge") + 
          xlab("Airlines") + ylab("Delays") 


```
```{r}
 ggplot(airline_statusinfo7, aes(x = Airline, y=ontime_rate, fill = City)) +  
          geom_bar(stat="identity",position="dodge") + 
          xlab("Airlines") + ylab("On time rate") 


```

## In conclusion

in conclusion we can witness on the chart how AM WEST	has slightly higher delay percentage than Alaska. Also we see how ontime rate is very similar/close for both airlines.  

