---
title: "DATA 606 Data Project Proposal"
author: "John Mazon"
---

### Data Preparation


```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(RCurl)
library(Hmisc)
library(ggplot2)

```


### Research question 

Is there a correlation in diamond value with respect to it's carat amount and cut quality?  


### Cases 

This dataset contains the prices and various attributes of almost 54,000 diamonds. Specifically, it is a data frame with 53940 rows and 10 variables. Some characteristics specified for each row are price[in U.S Dollars], carat, cut quality, color, clarity, total depth percentage and table.   



### Data collection 

This classic dataset has been found on the Kaggle site. 



### Type of study 

This is an observational study since we are trying to infer from already collected data.No specific experiment is being conducted data is being observed only.



### Data Source 

Original data is from Kaggle site:
https://www.kaggle.com/shivam2503/diamonds




### Dependent Variable

The response variable is the diamond value and it is quantitative.


### Independent Variable

Carat amount is the quantitative independent variable
Cut quality is the qualitative independent variable


### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}
my_data <- getURL("https://raw.githubusercontent.com/johnm1990/DATA607/master/diamonds.csv")
diamond_info <- read.csv(text=my_data)

diamond_df <- data.frame(diamond_info)


```

```{r}
summary(diamond_df$price)


```

```{r}

summary(diamond_df$carat)


```

```{r}

summary(diamond_df$cut)

```

```{r}
boxplot(diamond_df$price, main = "Price")

```

```{r}

boxplot(diamond_df$carat ~ diamond_df$cut)
```






