---
title: "Week 9 Assignment"
author: "John Mazon"
date: "10/24/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(jsonlite)
library(tidyverse)
library(httr)
library(stringr)
library(DT)

```

## Introduction
The New York Times web site provides a rich set of APIs, as described here:https://developer.nytimes.com/apis
We started by signing up for an API key.The task was to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R DataFrame.

#### ESTABLISH THE API PATH and KEY (PREPARE THE DATA)
Chose the topstories from the HTTP Request section of https://developer.nytimes.com/
Below we imported the apikey and path into respective objects
We created the initial dataframe 'magazine_df' 

```{r}


apikey <- "CwEonEZVnTIZET6h3EHtAKCZF1PmchR7"
url <- paste0("https://api.nytimes.com/svc/topstories/v2/health.json", "?api-key=", apikey)
magazine <- GET(url)
# names of the arrays from the API call

magazine_text <- content(magazine, "text")
magazine_df <- data.frame(fromJSON(magazine_text))


```
#### Workflow and Viewing Info of the data frame
Below we take a brief look at the column names currently of the dataframe

```{r}
(c_names <- names(magazine_df))

```

## TIDY/ORGANIZE THE DATA 
Below we tidy/organize the data by selecting the specific information we would like to display
We label our column names as desired to match the content
```{r}

FilteredData = magazine_df[, c("results.section", "results.title","results.abstract","results.url" )]
colnames(FilteredData) = c("Section","Title","Abstract","Link") 

colnames(FilteredData)
str(FilteredData)

```

Using 'datatable' function we pull in our final 'FilteredData' frame to organize the information to be displayed. We can also customize coloring of borders as well as alignment here, 
```{r}

datatable(FilteredData, colnames = c('Entry #', 'Section', 'Title', 'Abstract', 'Link'), class = 'cell-border stripe', options = list(
  initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#5a5a5a', 'color': '#fff', 'text-align': 'center !important'});",
    "$(this.api().table().body()).css({'color': '#000', 'text-align': 'center !important'});",
    "}")
))

```



## Conclusion
The initial dataframe created after the importing of the API path and url with key is heavily untidy. We need the help of a few tools in order to structure the data. Using library DT 'datatable' it makes it much easier to create a pleasant visual distribution of the data including search option
